### 2021-06-07 paper/report.tex
Resolve some comments.

We apply this model to Germany. In March and April 2020, the country broke the first wave
of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
period mid-September 2020 to the end of May 2020. We pick the starting date for two
reasons. First, we do not include the first wave because the environment was very
different (e.g., aggregate PCR test capacity was much lower and we would require a very
different model for calculating the share of known cases) and some data was not recorded
yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model national
and international travel which you mention below. Regarding data I think everything would
be available. The low numbers during summer in our sample size might also be a problem
because Covid might go extinct in our model when case numbers are so low.}. Second, a
large fraction cases during summer of 2020 were traced to international
travel,\comment[id=HM]{cite \url{https://www.rki.de/DE/Content/Infekt/EpidBull/Archiv/2021/Ausgaben/08_21.pdf}}, but the precise number is difficult to model.

Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the \replaced[id=HM]{color}{blue?} line in
Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
which tracks the tightness of non-pharmaceutical interventions. We transform the index so
that lower values represent higher levels of restrictions. A value of zero means all
measures incorporated in the index are turned on. The value 1 represents the situation of
mid-September, where half of all restrictions incorporated in the index are
active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
need the API access...}. In the six weeks between mid September and mid October, cases
increased by a factor of 3.\comment[id=Klara]{In the next iteration all empirical data is
in the figures/results/tables/empirical_analogues.csv}
% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
Restrictions were somewhat tightened in mid-October and again in early November. New
infections remained constant throughout November, before rising again in December.

@@

Quickly rising rates in December prompted the most stringent lockdown to this date. Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores. Shortly afterwards, more factors become relevant. Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just over 40\% had received at least one dose of a vaccine. Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies. At-home tests approved by authorities were ... in March.\comment[id=HM]{update precise timeline}\comment[id=Klara]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}). First tests to self-administer became available starting March 6. However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}).} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters. The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=Klara]{Will be part of the empirical analogues table}. In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public. In some cases, customers were only allowed to enter with a recent negative rapid test result.

These developments are characteristic of many countries: The initial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests. At broadly similar points in time, novel strains of the virus have started to pose additional challenges. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well. This is also true for infections by age and geographical region, which are shown in the supplementary materials (figures~\ref{fig:age_group_fit, fig:state_fit}. We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic. Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks. For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%. This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality. Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} for their individual contributions. Two things are particularly interesting. First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality. Second, the most stringent regulations are associated with the period of strong decreases in new infection beteen late December 2020 and mid-February 2021. The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period. The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts outside of work and school\comment[id=HM]{check with corrected graph}.

@@

The effect of the vaccination campaign is surprisingly small at first sight. The reasons for this are the slow start---10\% of the population was vaccinated in XX date, 20\% in mid-April\comment[id=HM]{(check!)}---and the focus on older individuals. These groups contribute less to the spread of the disease than others due to a lower number of contacts, see~\ref{fig:assortativity}. It is important to note that the initial focus of the campaign was to prevent deaths and severe disease; the case fatality was rate considerably lower during the third wave when compared to the second (4.4\% between October and February and 1.4\% between March and June).

>>>>

We apply this model to Germany. In March and April 2020, the country broke the first wave
of the pandemic fairly quickly. Between mid-May and mid-September, daily new infections
were below 20 per Million and day.\comment[id=HM]{Cite our world in data} We model the
period mid-September 2020 to the end of May 2020. We pick the starting date for two
reasons. First, we do not include the first wave because the environment was very
different (e.g., aggregate PCR test capacity was much lower and we would require a very
different model for calculating the share of known cases) and some data was not recorded
yet.\comment[id=HM]{True?}\comment[id=K]{We would definitely need to model national
and international travel which you mention below. Regarding data I think everything would
be available. The low numbers during summer in our sample size might also be a problem
because Covid might go extinct in our model when case numbers are so low.}. Second, a
large fraction cases during summer of 2020 were traced to international
travel \citep{KochInstitut2021}, but the precise number is difficult to model.

Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the \replaced[id=HM]{color}{blue?} line in
Figure~\ref{fig:stringency_index} the Oxford Response Stringency Index \citep{Hale2020},
which tracks the tightness of non-pharmaceutical interventions. We transform the index so
that lower values represent higher levels of restrictions. A value of zero means all
measures incorporated in the index are turned on. The value 1 represents the situation of
mid-September, where half of all restrictions incorporated in the index are
active.\comment[id=HM]{Need to find version of index with a couple of measures, probably
need the API access...}. In the six weeks between mid September and mid October, cases
increased by a factor of 3.\comment[id=K]{In the next iteration all empirical data is
in the figures/results/tables/empirical\_analogues.csv}
% the week of Sep 13 had an incidence of 12.4, the week of Oct 18 had 51.
Restrictions were somewhat tightened in mid-October and again in early November. New
infections remained constant throughout November, before rising again in December.

@@

Quickly rising rates in December prompted the most stringent lockdown to this date. Schools and daycare centers were closed again, so were customer-facing businesses except for grocery and drug stores. Shortly afterwards, more factors become relevant. Just after Christmas, the first people were vaccinated with a focus on older age groups and medical personnel (Figure~\ref{fig:antigen_tests_vaccinations}. By the end of May, just over 40\% had received at least one dose of a vaccine. Starting in January, rapid tests \dots These had to be administered by medical doctors or in pharmacies. At-home tests approved by authorities were ... in March.\comment[id=HM]{update precise timeline}\comment[id=K]{From a docstring:Bürgertests started in mid March but demand was very low initially (\url{https://bit.ly/3ehmGcj}). First tests to self-administer became available starting March 6. However, supply was very limited in the beginning (\url{https://bit.ly/3xJCIn8}). The first tests were licensed on Feb 24: \url{https://www.bundesregierung.de/breg-de/themen/coronavirus/zulassung-schnell-test-1861354}} From the peak of the second wave just before Christmas until the trough in mid-February, newly detected cases decreased by almost three quarters. The third wave in the spring of 2021 is associated with a rise in the share of the B.1.1.7 variant among all infections, depicted in Figure~\ref{fig:share_b117}, reaching more than 80\% in April\comment[id=HM]{Do we have the table somewhere?}\comment[id=K]{Will be part of the empirical analogues table}. In early March, some NPIs were being relaxed; e.g., hairdressers and home improvement stores were allowed to open again to the public. In some cases, customers were only allowed to enter with a recent negative rapid test result.

These developments are characteristic of many countries: The initial focus on NPIs to slow the spread of the disease has been accompanied by vaccines and a growing acceptance and use of rapid tests. At broadly similar points in time, novel strains of the virus have started to pose additional challenges. The blue line in Figure~\ref{fig:aggregated_fit} shows that our model tracks the simulated infections extremely well. This is also true for infections by age and geographical region, which are shown in the supplementary materials (figures~\ref{fig:age_group_fit, fig:state_fit}. We can disentangle various mechanisms due to the distinct temporal variation in the drivers of the pandemic. Next to the stringency index, the three lines in Figure~\ref{fig:stringency_index} summarize how contact reductions, increased hygiene regulations, and seasonality evolved since early September for each of the three broad contact networks. For example, a value of 0.75 for the work multiplier means that if the environment was the same as in September (levels of infection rates, no rapid tests or vaccinations, only the wildtype virus present), infections at the workplace would be reduced by 25\%. This reduction is the product of the effect of contact reductions, increased hygiene regulations, and seasonality. Along with the levels of infections, these measures completely determine the spread of SARS-CoV-2 in 2020.\comment[id=HM]{True?} See Appendix~\comment[id=HM]{Reference} \comment[id=K]{We have not done the decomposition for 2020 yet. Is on my todo list} for their individual contributions. Two things are particularly interesting. First, all lines broadly follow the stringency index and they would do so even more if we left out seasonality. Second, the most stringent regulations are associated with the period of strong decreases in new infection between late December 2020 and mid-February 2021. The measures were not enough, however, to stop the B.1.1.7 variant from spreading in the subsequent period. The steep drop in recorded cases during May 2021 is associated with rapid tests to X-Y~percent of the population, a vaccination rate that rose from 25\% to 4X\%, and mostly seasonality impacting a fall in the relative infectiousness of contacts outside of work and school\comment[id=HM]{check with corrected graph}.

@@

The effect of the vaccination campaign is surprisingly small at first sight. The reasons
for this are the slow start---10\% of the population had received their first vaccination
by the end of March, % 2021-03-24
20\% by April 19th ---and the focus on older individuals. These groups contribute less to
the spread of the disease than others due to a lower number of contacts,
see~\ref{fig:assortativity}. It is important to note that the initial focus of the
campaign was to prevent deaths and severe disease; the case fatality was rate
considerably lower during the third wave when compared to the second (4.4\% between
October and February and 1.4\% between March and June).
